Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-07-21 Gedoen.tex
redo v1"

\subsection{Chapter 1}
\begin{itemize}
	\item Het 1.4.21 reg gemaak. Moet nog deur gaan en seker maak ek het dit ander plekke reg.
\end{itemize}

\subsection{Chapter 2}
\begin{itemize}
	\item Het verbeteringe aangebring.
\end{itemize}

\subsection{Appendix}
\begin{itemize}
	\item $L_2$ in gesit.
\end{itemize}

>>>>

\section{Section 2.3}
\begin{itemize}
    \item[-] Page 42 - $(a+b)^2 \leq 2a^2 +2b^2$ verander na $(a+b)^2 < 2a^2 + 2b^2$
    \item[-] Page 42 - It follows that... was display, verander na in-line en ook die volgorde was nie dieselfde nie.
    \item[-] Page 42 - Hence het nie 'n komma gehad nie
\end{itemize}

\section{Section 2.4.2}
\begin{itemize}
    \item[-] Page 48 - Problem GVar is korrek vernader na die equation ref, die woord `satisfies' het wel weg geraak. Nou terug gesit.
\end{itemize}

### 2023-07-21 files/Chapter2/generalvibrationmodel.tex
redo v1"

The test functions $T[0,1]$ do not form a complete space. Let $V(0,1)$ be the closure of {$F_1(0,1)$} in $H^1(0,1)$ (see Section 1.4.4). It follows that $V(0,1)$ is complete and the product space {$V(0,1)\times V(0,1)$} denoted by $V$ is complete. It is called the energy space.\\

@@

\begin{proof}
	Let $f \in V$, then $f_1(0) = f_2(0) = 0$. Following from the corollary, $||f_1|| \leq ||f_1'||$ and $||f_2|| \leq ||f_2'||$. Therefore,
	\begin{eqnarray*}
		||f_1'|| & = & ||f_1 + f_2 - f_2||,\\
				 & \leq & ||f_1 + f_2|| + ||f_2||.\\
		||f_1'||^2 & \leq & 2||f_1 + f_2||^2 +||f_2'||^2.
	\end{eqnarray*}

	Hence from the definition of the norm $||\cdot||_X$, and again from the corollary,
	\begin{eqnarray*}
		||f||_X^2 & = & ||f_1||^2 + ||f_2||^2,\\
				  & \leq & 2||f_1 + f_2||^2 + 2||f_2'||^2.
	\end{eqnarray*}

	It follow from the inequality \eqref{C} that
	\begin{eqnarray}
		\Vert f \Vert_{W}^2 \leq 2C_2||f_1'-f_2||^2 + 2C_2||f_2'||^2. \label{f}
	\end{eqnarray}

@@

	Now combine the inequalities \eqref{f} and \eqref{b},
	\begin{eqnarray*}
		||f||_W^2 \leq Kb(f,f),
	\end{eqnarray*} for all $f \in V$.
\end{proof}

@@

Combining the solutions of the eigenvalue problem with the solution \eqref{eq:1D_Model:ModalAnalysisTn} yields:
\begin{eqnarray*}
	w_n(x,t) & = & T_n(t)\tilde{w}_n(t), \\
	\phi_n(x,t) & = & T_n(t)\tilde{\phi}_n(x).
\end{eqnarray*} 
Substitution show that these solutions does indeed satisfy the differential equations \eqref{eq:1D_Model:ModalPDE1} and \eqref{eq:1D_Model:ModalPDE2} with the boundary conditions. These solutions are the modal solutions and they are clearly periodic with natural angular frequencies $\sqrt{\lambda_n}$ (consequently the natural frequencies are $\frac{\sqrt{\lambda_n}}{2 \pi})$.

@@

\subsubsection*{Problem GVar}\label{sssec:existence:ProblemGVar}
Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
\begin{eqnarray}
c(u''(t),v)+b(u(t),v) = 0 \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
\end{eqnarray}
with initial conditions $u_{0}$ and $u_{1}$.

@@

*****************Voorstel*****************************

@@

Denote the partial sum $u^{N}(t) = \sum_{n=1}^{N} T_{n}(t)w_n$. Let
\begin{eqnarray*}
	u_0^{N} = \sum_{n=1}^{N} T_n(0) w_n \ \ \textrm{ and } \ \ u_{1}^{N} =\sum_{n=1}^{N} T'_n(0) w_n.
\end{eqnarray*}

Substitution shows that these partial sums with initial conditions $u^N(0) = u^N_0$ and $(u^N)'(0) = u^N_1$ are solutions to Problem GVar.

The authors then define an error function $u^E_N = u - u^N$. Since both $u$ and $u^N$ satisfies Problem GVar, so does this error function with the following the initial conditions, $u^E_N(0) = u_0 - u^N_0$ and $(u^E_N)'(0) = u_1 - u^N_1$.

Let $\mathcal{E}$ denote the energy of the error function $u^E_N$. Since $\mathcal{E}(t) = \mathcal{E}(0)$ for all $t>0$ it follows that, \[||u(t) -  u^N(t)||_V^2 + ||u'(t) - (u^N)'(t)||^2_W = ||u_0 - u^N_0||_V^2 + ||u_1 - u^N_1||_W^2. \]

\textcolor{red}{Nie seker wat ek hier moet doen met die Fourier coefficients nie.}

*****************************************************

\textcolor{red}{**************** OU WEERGAWE******************************}

The authors also prove that for all $t>0$, $\mathcal{E}(t) \leq \mathcal{E}(0)$.

Denote the partial sum $u^{N}(t) = \sum_{n=1}^{N} T_{n}(t)w_n$. Let
\begin{eqnarray*}
	u_0^{N} = \sum_{n=1}^{N} A^*_n w_n \ \ \textrm{ and } \ \ u_{1}^{N} =\sum_{n=1}^{N} B^*_n w_n
\end{eqnarray*} where $A^*_n = T_n(0)$ and $B^*_n = T'_n(0)$.

Substitution shows that these partial sums with initial conditions $u^N(0) = u^N_0$ and $(u^N)'(0) = u^N_1$ are solutions to Problem GVar.

The authors then define an error function $u^E_N = u - u^N$. Since both $u$ and $u^N$ satisfies Problem GVar, so does this error function with the following the initial conditions, $u^E_N(0) = u_0 - u^N_0$ and $(u^E_N)'(0) = u_1 - u^N_1$.

Let $\mathcal{E}$ denote the energy of the error function $u^E_N$. Since $\mathcal{E}(t) \leq \mathcal{E}(0)$ for all $t>0$ it follows that, \[||u(t) -  u^N(t)||_V^2 + ||u'(t) - (u^N)'(t)||^2_W \leq ||u_0 - u^N_0||_V^2 + ||u_1 - u^N_1||_W^2. \]

The authors of \cite{CVV18} prove that both $||u_0 - u^N_0||_V^2 \rightarrow 0$ and $||u_1 - u^N_1||_W^2 \rightarrow 0$ as $N \rightarrow \infty$. Therefore $\mathcal{E}(t) \rightarrow 0$ as $N \rightarrow \infty$.

\textcolor{red}{**********************************************}

It follows that the partial series solution converges to the solution $u$ as the initial conditions $u_0^N$ and $u_1^N$ converges to $u_0$ and $u_1$ respectively.

@@

In this dissertation, the stated objectives were to compare different models for use in applications. These comparisons are made in Chapter 4 and 6. Being able to express the solutions of the models as a formal series solution, enable us to compare the different models by only considering the eigenvalues and eigenfunctions of the different models.

The validity of the formal series solution with initial conditions is required to investigate the vibration of the models. This ensured that if the models are disturbed in the exact same way, the formal series solution remains valid, and a fair comparison can be made.

>>>>

The test functions $T[0,1]$ do not form a complete space. Let $V(0,1)$ be the closure of {$F_1(0,1)$} in $H^1(0,1)$ (see Section 1.4.4). It follows that $V(0,1)$ is complete and the product space {$V(0,1)\times V(0,1)$}, denoted by $V$, is complete. It is called the energy space.\\

@@

\begin{proof}
	Let $f \in V$, then $f_1(0) = f_2(0) = 0$. Following from the corollary, $||f_1|| \leq ||f_1'||$ and $||f_2|| \leq ||f_2'||$. Therefore,
	\begin{eqnarray*}
		||f_1'|| & = & ||f_1' - f_2 + f_2||,\\
				 & \leq & ||f_1' - f_2|| + ||f_2||,\\
				 & \leq & ||f_1' - f_2|| + ||f_2'||.
	\end{eqnarray*}

	It follows that $||f_1'||^2 \leq 2||f_1' - f_2||^2 + 2||f_2'||^2$, since $\displaystyle (a+b)^2 \leq 2a^2 + 2b^2$. Hence, from the definition of the norm $||\cdot||_X$, and again from the corollary,
	\begin{eqnarray*}
		||f||_X^2 & = & ||f_1||^2 + ||f_2||^2,\\
				  & \leq & ||f_1'||^2 + ||f_2'||^2,\\
				  & \leq & 2||f_1' - f_2||^2 + 3||f_2'||^2.
	\end{eqnarray*}

	It follows from the inequality \eqref{C} that
	\begin{eqnarray}
		\Vert f \Vert_{W}^2 \leq 2C_2||f_1'-f_2||^2 + 3C_2||f_2'||^2. \label{f}
	\end{eqnarray}

@@

	Now combine the inequalities \eqref{f} and \eqref{b}. There exists a positive constant $C_V$ such that
	\begin{eqnarray*}
		||f||_W^2 \leq C_V b(f,f),
	\end{eqnarray*} for all $f \in V$.
\end{proof}

@@

Combining the solutions of the eigenvalue problem with the solution \eqref{eq:1D_Model:ModalAnalysisTn} yields:
\begin{eqnarray*}
	w_n(x,t) & = & T_n(t)\tilde{w}_n(t), \\
	\phi_n(x,t) & = & T_n(t)\tilde{\phi}_n(x).
\end{eqnarray*} 
Substitution show that these solutions does indeed satisfy the differential equations \eqref{eq:1D_Model:ModalPDE1} and \eqref{eq:1D_Model:ModalPDE2} with the boundary conditions. These solutions are the modal solutions and they are clearly periodic with natural angular frequencies $\sqrt{\lambda_n}$ (and consequently the natural frequencies are $\frac{\sqrt{\lambda_n}}{2 \pi})$.

@@

\subsubsection*{Problem GVar}\label{sssec:existence:ProblemGVar}
Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
\begin{eqnarray}
c(u''(t),v)+b(u(t),v) = 0 \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
\end{eqnarray}
with initial conditions $u(0) = u_0$ and $u'(0) = u_1$.

@@

@@

Denote the partial sum $u^{N}(t) = \sum_{n=1}^{N} T_{n}(t)w_n$. Ideally
\begin{eqnarray*}
	u_0^{N} = \sum_{n=1}^{N} T_n(0) w_n \ \ \textrm{ and } \ \ u_{1}^{N} =\sum_{n=1}^{N} T'_n(0) w_n,
\end{eqnarray*}
but $T_n$ is not uniquely defined by the differential equation.

Substitution shows that these partial sums with initial conditions $u^N(0) = u^N_0$ and $(u^N)'(0) = u^N_1$ are solutions for \eqref{eq:existence:ProblemGVarHom} in Problem GVar.

The authors then define an error function $u^E_N = u - u^N$. Since both $u$ and $u^N$ satisfies \eqref{eq:existence:ProblemGVarHom}, so does this error function with the following the initial conditions, $u^E_N(0) = u_0 - u^N_0$ and $(u^E_N)'(0) = u_1 - u^N_1$.

Let $\mathcal{E}$ denote the energy of the error function $u^E_N$. Since $\mathcal{E}(t) = \mathcal{E}(0)$ for all $t>0$ it follows that, 
\begin{eqnarray}
	||u(t) -  u^N(t)||_V^2 + ||u'(t) - (u^N)'(t)||^2_W = ||u_0 - u^N_0||_V^2 + ||u_1 - u^N_1||_W^2. \label{eq:inequality}
\end{eqnarray}

Now, $u_0$ and $u_1$ are given in Problem GVar. Therefore the generalized Fourier coefficients for $u_0$ and $u_1$ must be used to compute $u_0^N$ and $u_1^N$. These Fourier coefficients are $\displaystyle u_0^N = \sum_{n=1}^{N} b(u_0, w_n)w_n$ and $\displaystyle u_1^N = \sum_{n=1}^{N} c(u_1, w_n)w_n$.

Then $||u_0 - u^N_0||_V^2 \rightarrow 0$ and $||u_1 - u^N_1||_W^2 \rightarrow 0$ as $N \rightarrow \infty$. Therefore $\mathcal{E}(t) \rightarrow 0$ as $N \rightarrow \infty$ by \eqref{eq:inequality}.

It follows that the partial sums of the series solution converges to the solution $u$ as the initial conditions $u_0^N$ and $u_1^N$ converges to $u_0$ and $u_1$ respectively.

@@

In this dissertation, the stated objectives were to compare different models for use in applications. These comparisons are made in Chapter 4 and 6. Being able to express the solutions of the models as a formal series solution, enable us to compare the different models by only considering the eigenvalues and eigenfunctions of the different models. This is explained in detail in \cite{CVV18}.

The validity of the formal series solution with initial conditions is required to investigate the vibration of the models. This ensured that if the models are disturbed in the exact same way, the formal series solution remains valid, and a comparison can be made.

### 2023-07-21 files/Chapter3/dynamicconvergence.tex
redo v1"

\documentclass[../../main.tex]{subfiles}
\insubfile{\AtEndDocument{\bibliography{../../references.bib}}}

\begin{document}
\section{Introduction}
In this chapter, two important convergence results of the Finite Element Method
(FEM) is discussed. The first result looks at the convergence of the Galerkin
approximation for second order hyperbolic type problems, as presented
in~\cite{BV13}. The second result, from the textbook~\cite{SF97}, examines the
convergence of eigenvalues and eigenfunctions for a one-dimensional vibration
problem, using the Finite Element Method.

@@

\subsection{Galerkin approximation of Problem GVar}\label{subsec:galerkin-approximation-of-problem-gvar}
Recall from Chapter 2 that $V$ is the closure of the test function space $T$ in
the Hilbert space $H$.

Suppose $S^h$ is a finite n-dimensional subspace of V. Let $\{\phi_1, \
	\phi_2,...,\phi_n\}$ be a set of linear independent basis functions such that
$S^h = \textrm{span}(\{\phi_1, \phi_2, ..., \phi_n \})$ and $\phi_i \in
	C^{(1)}$ for all $i$.

For any function $w_h$ in $S^h$, $w_h$ can be expressed as a linear combination
\begin{eqnarray*}
	w_h = \sum_{j=1}^n w_j \phi_j.
\end{eqnarray*} Define a vector valued function $\bar{w} = \langle w_1,w_2,...,w_n \rangle$ for each $w_h \in S^h$. The problem of solving the Galerkin Approximation is denoted by Problem GVar$^h$.

@@

\begin{eqnarray}
	e^h(t) = u(t) - u^h(t). \label{Error}
\end{eqnarray}

In \cite{BV13}, a projection $P$ is defined so that $Pu \in V$. The error is
then split using this projection. Let $e(t) = Pu(t)- u_h(t)$ and $e_p(t) = u(t)
	- Pu(t)$ so that
\begin{eqnarray}
	e^h(t) = e_p(t) +e(t) \label{Error}
\end{eqnarray}

The authors of \cite{BV13} mention that interpolation theory can be used to determine the error $e_p$. However to obtain an error estimate for $e$ is not as trivial. This is where the authors use the Finite Element Analysis, which is used to obtain an error estimate for $e$.

@@

Using the definition of the error $e^h$ while also adding and subtracting
$Pu(t)$ in the bilinear form $b$, we can write
\begin{eqnarray*}
	c(e''(t),v) + a(e'(t),v) + b(u(t) - Pu(t),v) + b(Pu(t) +u_h(t),v)- = 0
\end{eqnarray*}

@@

In \cite{BV13}, this is presented as a proposition.
\newtheorem{DC_Prop1}{Proposition}
\begin{DC_Prop1}
	If $u$ is the solution of Problem GVar and $u$ satisfies assumption \textbf{B1}, then
	\begin{eqnarray}
		c(e''(t),v) + a(e'(t),v) + b(e(t),v) = 0 \ \ \ \textrm{ for all } v \in V.
	\end{eqnarray}
\end{DC_Prop1}

@@

\newtheorem{col_1}{Corollary}
\begin{col_1}
	If $u$ is the solution of Problem GVar with $a = 0$ and $u$ satisfies assumption \textbf{B1}, then
	\begin{eqnarray}
		\Vert e(t)\Vert_{W} \leq \sqrt{2} \Vert e(0) \Vert_W + 2T||e'_h(0)||_W + 4\sqrt{T}\max_{t \in [0,T]} \Vert e'_p(t) \Vert_W.
	\end{eqnarray}
\end{col_1}

@@

The error $e(t)$, can now be expressed as
\begin{eqnarray}
	e(t_k) = u(t_k) - u^h_k & = & [u(t_k)-u_h(t_k)] + [u_h(t_k) - u^h_k]. \label{error_e}
\end{eqnarray}

@@

\subsubsection{Notation}
For any sequence $\{x_k\} \in R_n$,
\begin{eqnarray*}
	\delta_{t}x_{k}& = &\tau^{-1}[x_{k+1}-x_{k}],\\
	x_{k+\frac{1}{2}} & = & \frac{1}{2}[x_{k+1}+x_{k}].
\end{eqnarray*}

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, two important convergence results of the Finite Element Method
(FEM) is discussed. The first result looks at the convergence of the Galerkin
approximation for second order hyperbolic type problems, as presented
in~\cite{BV13}. The second result, from the textbook~\cite{SF97}, examines the
convergence of eigenvalues and eigenfunctions for a one-dimensional vibration
problem, using the Finite Element Method.

@@

Recall from Chapter 2 that $V$ is the closure of the test function space $T$ in
the Hilbert space $H$.

@@

Let $u$ be the solution of Problem GVar and $u^h$ be the solution of Problem GVar$^h$.
 The authors of \cite{BV13} define the following error,
\begin{eqnarray}
	e^h(t) = u(t) - u^h(t). \label{Error}
\end{eqnarray}

In \cite{BV13}, a projection $P$ is defined so that $Pu \in V$. The error is
then split using this projection. Let $e(t) = Pu(t)- u_h(t)$ and $e_p(t) = u(t)
	- Pu(t)$ so that
\begin{eqnarray}
	e^h(t) = e_p(t) +e(t). \label{Error}
\end{eqnarray}

According to the authors of~\cite{BV13}, interpolation theory can be used to determine the error $e_p$. However to obtain an error estimate for $e$ is not as trivial. This is where the use of the Finite Element Analysis comes in.

To simplify the proof, the problem is divided into a semi-discrete problem and a fully discrete problem.

@@

Using the definition of the error $e^h$ while also adding and subtracting
$Pu(t)$ in the bilinear form $b$, we can write
\begin{eqnarray*}
	c(e_h''(t),v) + a(e_h'(t),v) + b(u(t) - Pu(t),v) + b(Pu(t) +u_h(t),v)- = 0.
\end{eqnarray*}

@@

This equation \eqref{DC_E3} is presented in \cite{BV13} as a proposition.
\newtheorem{DC_Prop1}{Proposition}
\begin{DC_Prop1}
	If $u$ is the solution of Problem GVar and $u$ satisfies assumption \textbf{B1}, then
	\begin{eqnarray}
		c(e_h''(t),v) + a(e_h'(t),v) + b(e(t),v) = 0 \ \ \ \textrm{ for all } v \in V.
	\end{eqnarray}
\end{DC_Prop1}

@@

\newtheorem{col_1}{Corollary}
\begin{col_1}
	If $u$ is the solution of Problem GVar with $a = 0$ and $u$ satisfies assumption \textbf{B1}, then
	\begin{eqnarray}
		\Vert e_(t)\Vert_{W} \leq \sqrt{2} \Vert e(0) \Vert_W + 2T||e'_h(0)||_W + 4\sqrt{T}\max_{t \in [0,T]} \Vert e'_p(t) \Vert_W.
	\end{eqnarray}
\end{col_1}

\begin{proof}
	Assume that there is no damping other than the weak damping, i.e. $a = 0$. The the equation in proposition 1 becomes
	\begin{eqnarray*}
		c(e_h''(t),v) + b(e(t),v) = 0 \ \ \ \textrm{ for all } v \in V.
	\end{eqnarray*}

	If $v \in S^h[0,T]$, then 
	\begin{eqnarray*}
		c(e_h''(t),v(t)) + b(e(t),v(t)) = 0.
	\end{eqnarray*}

	Let $v'(t) = c(t)$, then 
	\begin{flalign}
		\frac{d}{dt}\left[\frac{1}{2}c(e,e)-\frac{1}{2}b(v,v)-c(e_h,v) \right] &= c(e',e) - b(e,v)- c(e''_h, v)-c(e_h',v')\nonumber\\
		&= c(e'_p,e). \label{DC_E4}
	\end{flalign}

	Choose $v(\tau) = 0$ for some $\tau \in [0,T]$, and integrate \eqref{DC_E4} between $0$ and $\tau$ to obtain
	\begin{eqnarray*}
		a
	\end{eqnarray*}

\end{proof}

@@

The error $e(t)$, can now be expressed as
\begin{eqnarray}
	e(t_k) = u(t_k) - u^h_k & = & [u(t_k)-u_h(t_k)] + [u_h(t_k) - u^h_k]. \label{error_ek}
\end{eqnarray}

@@

\subsubsection{Notation}
For any sequence $\{x_k\} \in R_n$,
For any sequence $\{x_k\} \in R_n$,
\begin{eqnarray*}
	\delta_{t}x_{k}& = &\tau^{-1}[x_{k+1}-x_{k}],\\
	x_{k+\frac{1}{2}} & = & \frac{1}{2}[x_{k+1}+x_{k}].
\end{eqnarray*}
